Varieties of representable relation algebra reducts

### 2023-02-05 varieties-reducts/Text.tex
residuated BAOs

@@

\begin{definition}
    A proper relation algebra is an algebra $\mathcal{R} = (R, \emptyset,W, \cup, -, |, {}^{\smile}, {\bf 1 })$ such that $R \subseteq \mathcal{P}(W)$, where $W \subseteq X \times X$ is an equivalence relation; $|$ is relation composition, ${}^{\smile}$ is relation converse, ${\bf Id}$ is a
    diagonal subset of $W$, that is:
    \begin{enumerate}
    \item $a | b = \{ (x, z) \: | \: \exists y \: (x, y) \in a \: \& \: (y, z) \in b \}$
    \item $a^{\smile} = \{ (x, y) \: | \: (y, x) \in a \}$
    \item ${\bf Id} = \{ (x, y) \: | \: x = y \}$
    \end{enumerate}
      The class of all proper relation algebras is denoted as ${\bf PRA}$. ${\bf Rs}$ is the class of all relation set algebras, proper relation algebra with a diagonal subrelation as an identity. ${\bf RRA}$ is the class of all representable relation algebras, that is, the closure of ${\bf PRA}$ under isomorphic copies. That is,
      ${\bf RRA} = {\bf IPRA}$.
\end{definition}

\subsection{Varieties}
\subsection{Non-varieties}
\subsection{Unknown}

>>>>

\section{Question}

Let $\tau$ be a subsignature of operations expressible in the signature of $RA$. ${\bf R}(\tau)$ is a variety iff ??.

Investigate discriminators for representable relation algebra reducts.

@@

\begin{definition}
    A proper relation algebra is an algebra $\mathcal{R} = (R, \emptyset,W, \cup, -, |, {}^{\smile}, {\bf 1 })$ such that $R \subseteq \mathcal{P}(W)$, where $W \subseteq X \times X$ is an equivalence relation; $|$ is relation composition, ${}^{\smile}$ is relation converse, ${\bf Id}$ is a
    diagonal subset of $W$, that is:
    \begin{enumerate}
    \item $a | b = \{ (x, z) \: | \: \exists y \: (x, y) \in a \: \& \: (y, z) \in b \}$
    \item $a^{\smile} = \{ (x, y) \: | \: (y, x) \in a \}$
    \item ${\bf Id} = \{ (x, y) \: | \: x = y \}$
    \end{enumerate}
      The class of all proper relation algebras is denoted as ${\bf PRA}$. ${\bf Rs}$ is the class of all relation set algebras, proper relation algebra with a diagonal subrelation as an identity. ${\bf RRA}$ is the class of all representable relation algebras, that is, the closure of ${\bf PRA}$ under isomorphic copies.
\end{definition}

\subsection{Varieties}

\subsubsection{Boolean algebras with residuated operators}

Let $\mathcal{A}_0$ be a Boolean algebra, a unary operation on $f$ is called \emph{residuated} if there exists a \emph{residual} operation $g$ such that for all $a, b \in \mathcal{A}_0$:
\begin{center}
$f(a) \leq b$ iff $a \leq f(b)$
\end{center}
Equivalently, $f$ is residuated if there exists a conjugate operation $h$ such that for all $a, b \in \mathcal{A}_0$:
\begin{center}
$f(a) \cdot b = 0$ iff $a \land h(b) = 0$
\end{center}
$f$ is \emph{self-conjugate} whenever $f$ is equal to its conjugate operation.

\begin{theorem}
Let $f$ be a residuated operator on $\mathcal{A}_0$, then

\begin{enumerate}
\item $f$ is normal and completely additive,
\item $f$ and $h$ are conjugate operations on $\mathcal{A}_0$ iff they are normal and the following holds for all $a, b \in \mathcal{A}_0$
\begin{center}
$f(a) \cdot b \leq f(a \cdot h(b))$ and $a \cdot h(b) \leq h(f(a) \cdot b)$
\end{center}
\end{enumerate}
\end{theorem}
Two $n$-ary operators $f$ and $h$ are conjugate in the $i$-th argument if $f_{\vec{a}, i}$ is conjugate to $g_{\vec{a}, i}$. Let $\mathcal{A}$ be a BAO, $\mathcal{A}$ is called a \emph{Boolean algebra with residuated operators} if for each operator $f$ of $\mathcal{A}$ and for each $i < ar(f)$ there exists an $ar(f)$-ry term $t$, which is conjugate to $f$ in the $i$-th argument. ${\bf RA}$ is an example, where $\smile$ is self-conjugate and composition has conjugate terms $x ; y^{\smile}$ and $x^{\smile} ; y$.

An $r$-algebra is a Boolean algebra with three residuated binary operations $\bullet$, $\triangleright$ and $\triangleleft$, where $\triangleright$ and $\triangleleft$ are right and left conjugates to $\bullet$, that is, the following conditions are equivalent:
\begin{enumerate}
\item $(x \bullet y) \cdot z = 0$
\item $(x \triangleright z) \cdot y = 0$
\item $(z \triangleleft y) \cdot x = 0$
\end{enumerate}

\subsubsection{Residuated monoids}

\subsubsection{Positive relation algebras}

\subsubsection{Domain algebras}

\subsection{Non-varieties}
\subsection{Unknown}
